---
title: "Assignment 3: Data Exploration"
author: "Yinsu Wang, Section #3"
output: pdf_document
geometry: margin=2.54cm
---

## OVERVIEW

This exercise accompanies the lessons in Environmental Data Analytics on Data Exploration. 

## Directions
1. Change "Student Name, Section #" on line 3 (above) with your name and section number.
2. Work through the steps, **creating code and output** that fulfill each instruction.
3. Be sure to **answer the questions** in this assignment document.
4. When you have completed the assignment, **Knit** the text and code into a single PDF file.
5. After Knitting, submit the completed exercise (PDF file) to the dropbox in Sakai. Add your last name into the file name (e.g., "FirstLast_A03_DataExploration.Rmd") prior to submission.

The completed exercise is due on <>.


## Set up your R session

1. Check your working directory, load necessary packages (tidyverse), and upload two datasets: the ECOTOX neonicotinoid dataset (ECOTOX_Neonicotinoids_Insects_raw.csv) and the Niwot Ridge NEON dataset for litter and woody debris (NEON_NIWO_Litter_massdata_2018-08_raw.csv). Name these datasets "Neonics" and "Litter", respectively. **Be sure to add the `stringsAsFactors = TRUE` parameter to the function when reading in the CSV files.**
```{r,  message = FALSE}
getwd()
setwd("/Users/wangyinsu/Desktop/2022 Spring/Env872/Environmental_Data_Analytics_2022")
library(tidyverse)
Neonics<-read.csv("/Users/wangyinsu/Desktop/2022 Spring/Env872/Environmental_Data_Analytics_2022/Data/Raw/ECOTOX_Neonicotinoids_Insects_raw.csv",stringsAsFactors = TRUE)
Litter<-read.csv("/Users/wangyinsu/Desktop/2022 Spring/Env872/Environmental_Data_Analytics_2022/Data/Raw/NEON_NIWO_Litter_massdata_2018-08_raw.csv",stringsAsFactors = TRUE)
```

## Learn about your system

2. The neonicotinoid dataset was collected from the Environmental Protection Agency's ECOTOX Knowledgebase, a database for ecotoxicology research. Neonicotinoids are a class of insecticides used widely in agriculture. The dataset that has been pulled includes all studies published on insects. Why might we be interested in the ecotoxicologoy of neonicotinoids on insects? Feel free to do a brief internet search if you feel you need more background information. 

> Answer: Probably because neonicotinoids have impacts not only on pests such as aphids, but also have negative impacts on insects such as bees, who are the most economically important group of pollinators worldwide; we thus need to find out ways to kill pests at same time to protect these good insects such as bees therefore we need to study the ecotoxicology of neonicotinoids. 

3. The Niwot Ridge litter and woody debris dataset was collected from the National Ecological Observatory Network, which collectively includes 81 aquatic and terrestrial sites across 20 ecoclimatic domains. 32 of these sites sample forest litter and woody debris, and we will focus on the Niwot Ridge long-term ecological research (LTER) station in Colorado. Why might we be interested in studying litter and woody debris that falls to the ground in forests? Feel free to do a brief internet search if you feel you need more background information. 

> Answer: this may be because some litter and trees that fall and decay in the forest add nutrients to the forest soil and retain moisture in the forest. Also they have a role in carbon budgets and nutrient cycling. 

4. How is litter and woody debris sampled as part of the NEON network? Read the NEON_Litterfall_UserGuide.pdf document to learn more. List three pieces of salient information about the sampling methods here: 

> Answer: 
*Spatial Sampling: Litter and fine woody debris sampling is executed at terrestrial NEON sites that contain woody vegetation >2m tall. Along with most of NEON’s plant productivity measurements, sampling for this product occurs only in tower plots. Locations of tower plots are selected randomly within the 90% flux footprint of the primary and secondary airsheds (and additional areas in close proximity to the airshed, as necessary to accommodate sufficient spacing between plots). 
* Spatial Sampling:In sites with forested tower airsheds, the litter sampling is targeted to take place in 20 40m x 40m plots. In sites with low-statured vegetation over the tower airsheds, litter sampling is targeted to take place in 4 40m x 40m tower plots (to accommodate co-located soil sampling) plus 26 20m x 20m plots.
*Spatial Sampling:Trap placement within plots may be either targeted or randomized, depending on the vegetation. In sites with > 50% aerial cover of woody vegetation >2m in height, placement of litter traps is random and utilizes the randomized list of grid cell locations being utilized for herbaceous clip harvest and bryophyte sampling.



## Obtain basic summaries of your data (Neonics)

5. What are the dimensions of the dataset?

```{r}
dim(Neonics)
```

6. Using the `summary` function on the "Effect" column, determine the most common effects that are studied. Why might these effects specifically be of interest?
```{r}
summary(Neonics$Effect)
```

> Answer: the most common effects that are studied are mortality and population. The reason why these two effects might be of interest is because mortality and population are the most important parameters to a species; they will finally influence a species's survival on Earth. 

7. Using the `summary` function, determine the six most commonly studied species in the dataset (common name). What do these species have in common, and why might they be of interest over other insects? Feel free to do a brief internet search for more information if needed.

```{r}
summary(Neonics$Species.Common.Name)
```

> Answer: the six most commonly studied species are:Honey Bee, Parasitic Wasp,Buff Tailed Bumblebee, Carniolan Honey Bee, Bumble Bee, Italian Honeybee.
          they all belongs to the suborder Apocrita. why they might be of interest over other insects might be that they are the most economically important group of pollinators worldwide and sensitive to pesticides such as neonicotinoids. 

8. Concentrations are always a numeric value. What is the class of Conc.1..Author. in the dataset, and why is it not numeric?

```{r}
class(Neonics$Conc.1..Author.)
```

> Answer:the class of this variable is "factor". The reason why it is not numeric is that this variable is used to categorize and store the data as integers, having a limited number of different values. 

## Explore your data graphically (Neonics)

9. Using `geom_freqpoly`, generate a plot of the number of studies conducted by publication year. 
```{r}
library(ggplot2)
ggplot(Neonics)+
  geom_freqpoly(aes(x=Publication.Year))
```

10. Reproduce the same graph but now add a color aesthetic so that different Test.Location are displayed as different colors.
```{r}
library(ggplot2)
ggplot(Neonics,aes(Publication.Year,color=Test.Location))+
  geom_freqpoly()
```

Interpret this graph. What are the most common test locations, and do they differ over time?

> Answer: As the publication year moves on, the number of studies increases first and then decreases. We also can conclude from this graph that the most common test location is lab, followed by field natural and field artifical; field underterminable is the least. As already said, all of these test locations differ over time, with firstly increase and then decrease. 

11. Create a bar graph of Endpoint counts. What are the two most common end points, and how are they defined? Consult the ECOTOX_CodeAppendix for more information. 

```{r}
ggplot(Neonics, aes(x = Endpoint)) +
  geom_bar()+
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
```

> Answer: the two most common endpoints are LOEL and NOEL. NOEL is defined as "No-observable-effect-level: highest dose (concentration) producing effects not significantly different from responses of controls according to author's reported statistical test (NOEAL/NOEC)" and LOEL is defined as "Lowest-observable-effect-level: lowest dose (concentration) producing effects that were significantly different (as reported by authors) from responses of controls (LOEAL/LOEC)". 
>

## Explore your data (Litter)

12. Determine the class of collectDate. Is it a date? If not, change to a date and confirm the new class of the variable. Using the `unique` function, determine which dates litter was sampled in August 2018. 
```{r}
library(lubridate)
class(Litter$collectDate)#it is "factor", not date.
Litter$collectDate<-ymd(Litter$collectDate) 
class(Litter$collectDate)
unique(Litter$collectDate)# litter was sampled on 2018-08-02 and 2018-08-30.
```

13. Using the `unique` function, determine how many plots were sampled at Niwot Ridge. How is the information obtained from `unique` different from that obtained from `summary`?

```{r}
unique(Litter$plotID)
summary(Litter$plotID)
```

> Answer: There are 12 plots sampled at Niwot Ridge. The information obtained from "summary" contains frequency of each variable but information from "unique" does not. 

14. Create a bar graph of functionalGroup counts. This shows you what type of litter is collected at the Niwot Ridge sites. Notice that litter types are fairly equally distributed across the Niwot Ridge sites.
```{r}
ggplot(Litter, aes(x = functionalGroup)) +
  geom_bar()
```

15. Using `geom_boxplot` and `geom_violin`, create a boxplot and a violin plot of dryMass by functionalGroup. 
```{r}
ggplot(Litter, aes(y = dryMass,x=functionalGroup,color=functionalGroup)) +
  geom_boxplot()
ggplot(Litter, aes(y = dryMass,x=functionalGroup,color=functionalGroup)) +
  geom_violin(draw_quantiles = c(0.25, 0.5, 0.75), scale = "width")
```

Why is the boxplot a more effective visualization option than the violin plot in this case?

> Answer: While a box plot only shows summary statistics such as median and interquartile ranges, the violin plot shows the full distribution of the data. In this case, boxplot can easily shows the quantiles of data and the outliers but in violin plot we can see the whole distribution of the variable, which makes it complicated.

What type(s) of litter tend to have the highest biomass at these sites? 

> Answer: Needles.